---
title: "Bash shebang"
author: "Nicholas Hathaway"
---

```{r setup, echo=FALSE}
library(knitr)
opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = TRUE)
options(width = 80)

```

When writing a plain text file for any various langauges like python, bash, R, etc. you can call the interperter directly on the file like below.

***a_python_script.py***
```{r, engine='python', eval = F}
import sys
def hello(name):
  sys.stdout.write("hello" +  name + "\n")

name = "nick"
hello(name)

```

```{r, engine='bash', eval = F}
python a_python_script.py
```

```{r, engine='python', eval = T, echo =F}
import sys
def hello(name):
  sys.stdout.write("hello " +  name + "\n")

name = "nick"
hello(name)
```

***a_R_script.R***
```{r, eval = F}
print("hello world")
```

```{r, engine='bash', eval = F}
Rscript a_R_script.R
```

```{r, eval = T, echo =F}
print("hello world")
```

Or you can make the file executable using chmod and then add a line to the very top of your file called the [she-bang](https://en.wikipedia.org/wiki/Shebang_(Unix)) line which supply the interperter. 

Add she-bang line  
***a_python_script.py***
```{r, engine='python', eval = F}
#!/usr/bin/python
import sys
def hello(name):
  sys.stdout.write("hello" +  name + "\n")

name = "nick"
hello(name)

```
You have to also make the file executable, if you don't you'll see this line

```{r, engine = "bash", eval = F}
./a_python_script.py
```

```{r, engine = 'bash', echo =F}
echo "-bash: ./a_python_script.py: Permission denied"
```

Make file executable, just has to be done one time
```{r, engine='bash', eval = F}
chmod +x a_python_script.py
```

Now you can just run the file and it will be interpreted by python
```{r, engine='bash', eval = F}
./a_python_script.py
```

```{r, engine='python', eval = T, echo =F}
import sys
def hello(name):
  sys.stdout.write("hello " +  name + "\n")

name = "nick"
hello(name)
```

The she-bang always starts with `#!` and should be on the very first line of the file. What fallows the `#!` is the path to the program you want to have interpret the file. The above with giving the path `/usr/bin/python` will probably work for most default systems but might not work on things like a large computer cluster where you have use `lsf` and `module load`. So a better way to get the right interperter is to use the program `env` which will search your environment for the interperter that you give it ensuring the right interperter is found. 

***a_python_script.py***
```{r, engine='python', eval = F}
#!/usr/bin/env python
import sys
def hello(name):
  sys.stdout.write("hello" +  name + "\n")

name = "nick"
hello(name)

```

So instead of giving the direct path to the python interperter you can query your environment for `python` and `env` will look at your `PATH` to find the correct one, this will also help to find programs if you have them instead in non-default places. Also using the `/usr/bin/env python` will make it easier to port your code around, giving it a higher chance of working when you use it on default machines and setups. 

You can do the same for any program like Rscript as well 

***a_R_script.R***
```{r, eval = F}
#!/usr/bin/env Rscript
print("hello world")
```





