---
title: "Downloading 2bit Genome Files"
author: "Nicholas Hathaway"
---

```{r setup, echo=FALSE}
library(knitr)
opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = TRUE)
options(width = 120)

newPath = Sys.getenv("PATH")
newPath = paste("/usr/local/bin:/home/hathawan/hathaway/bin:", newPath,sep = "")
Sys.setenv(PATH=newPath)

```

#Downloading 2bit genomes

The python library for ucscgenome can be used to download genome

```{r, engine='bash', eval = F}
pip install ucscgenome
#for local install on servers use --user if you don't have rights to install globaly
pip install --user ucscgenome

```


Then within python you can download by creating a genome object with the name of the genome you want, which will automatically donwload the genome to `~/.ucscgenome`

For example below would download the human genome version 19 and 38 
```{r , engine='python'}
#!/usr/bin/env python

from ucscgenome import Genome

hg19 = Genome('hg19')

hg38 = Genome('hg38')

```


```{r, engine='bash', eval = T}
ls ~/.ucscgenome
```


Go to here http://hgdownload.cse.ucsc.edu/downloads.html to see a list of available genomes


#Reading 2bit genomes
To read the downloaded 2bit files you can use the python package twobitreader
```{r, engine='bash', eval = F}
pip install twobitreader
#for local install on servers use --user if you don't have rights to install globaly
pip install --user twobitreader

```

Get the sequence of one of the chromosomes
```{r , engine='python'}
#!/usr/bin/env python

import twobitreader,os
from os.path import expanduser
home = expanduser("~")

hg38_genome = twobitreader.TwoBitFile(os.path.join(home, ".ucscgenome/hg38.2bit") )
chr9 = hg38_genome["chr9"]
print(chr9.get_slice(133255319, 133255568))

```


```{r , engine='python', eval = F, echo =F}
#!/usr/bin/env python

import twobitreader
from os.path import expanduser
home = expanduser("~")

with open("hg38.fasta", "w") as output:
    twobit = twobitreader.TwoBitFile(os.path.join(home, ".ucscgenome/hg38.2bit") )
    twobitreader.twobit_reader(twobit,write=output)
```



