---
title: "Get SRA run ids"
author: "Nicholas Hathaway"
---

```{r setup, echo=FALSE}
library(knitr)
opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = F)
options(width = 120)
newPath = Sys.getenv("PATH")
newPath = paste("/usr/local/bin/:", newPath,sep = "")
Sys.setenv(PATH=newPath)
```

#Querying SRA site for info

The following script contains code to query SRA for SRA numbers from expeirment(SRX/ERX), sample(SRS/ERS), project(SRP/XRP), and BioProject (PRJNA) identifiers, here's a link to download script here [getSraRunsFromAccIds.py](scripts/python/getSraRunsFromAccIds.py)

The script contains a class with static methods to get the RUN ids associated with other identifiers and a function to automatically generate the URL for the run ID. The script can be included in another script to use the SRAUtils class and it's functions or the script can be ran itself to get the ids and urls for several IDs

The script depends on the non-standard `requests` library which might need to be installed

```{r, engine='bash'}
pip install requests
```

***getSraRunsFromAccIds.py***
```{r , engine='python', eval = F}
#!/usr/bin/env python
import shutil, os, argparse, sys, stat
import requests
import csv, io
import traceback

class SRAUtils:
    '''
    Prefixes can be found here http://www.ddbj.nig.ac.jp/sub/prefix.html for DDBJ, ENA/EBI (ERA), and NCBI(SRA)
    '''
    @staticmethod
    def getInfoTableFromSearchTerm(search):
        payload = {"save": "efetch","db": "sra","rettype" : "runinfo", "term" : search };
        r = requests.get('http://trace.ncbi.nlm.nih.gov/Traces/sra/sra.cgi', params=payload)
        if 200 ==  r.status_code:
            if r.text.isspace():
                raise Exception("Got blank string from " + str(r.url ))
            else:
                reader_list = csv.DictReader(io.StringIO(r.text))
                infoRows = []
                for row in reader_list:
                    infoRows.append(row)
                if 0 == len(infoRows):
                    raise Exception('Found %d entries in SRA for "%s" when expecting at least 1' % (len(infoRows), search))
                else:        
                    return infoRows
                return infoRows
        else:
            raise Exception("Error in downloading from " + str(r.url) + " got response code " + str(r.status_code))
        
    @staticmethod
    def getRunAccsFromInfoTable(infoTab):
        runInfo = []
        for row in infoTab:
            runInfo.append(str(row.get('Run')))
        return runInfo
            
        
    @staticmethod
    def getSraUrlFromRunAccession(accesion):
        if not type(accesion) is str:
            raise Exception("Error in getSraUrlFromRunAccession: accesion should be str, not " + str(type(accesion)))
        if len(accesion) < 7:
            raise Exception("Error in getSraUrlFromRunAccession: accession should be least 7 character long, not " + str(len(accesion)) + ", for " + str(accesion))
        
        if not accesion.startswith("ERR") and not accesion.startswith("SRR"):
            raise Exception("Error in getSraUrlFromRunAccession: accession should start with either ERR or SRR, not " + str(accesion[0:3]) )
        
        template = "ftp://ftp-trace.ncbi.nlm.nih.gov/sra/sra-instant/reads/ByRun/sra/{PREFIX}/{PREFIX}{PREFIXNUMS}/{ACCESION}/{ACCESION}.sra"
        return template.format(PREFIX = accesion[0:3], PREFIXNUMS = accesion[3:6], ACCESION = accesion)
    
    @staticmethod
    def getInfoFromRunAcc(run):
        if not run.startswith("ERR") and not run.startswith("SRR"):
            raise Exception("run should start with ERR or SRR, not: " + run)
        infoTab = SRAUtils.getInfoTableFromSearchTerm(run)
        runInfo = []
        for row in infoTab:
            if run == str(row.get('Run')):
                runInfo.append(row)
        if 0 == len(runInfo):
            raise Exception('Found %d entries in SRA for "%s" when expecting at least 1' % (len(runInfo), run))
        else:
            return runInfo

    @staticmethod
    def getInfoFromSubmissionAcc(submission):
        if not submission.startswith("ERA") and not submission.startswith("SRA"):
            raise Exception("submission should start with ERR or SRR, not: " + submission)
        infoTab = SRAUtils.getInfoTableFromSearchTerm(submission)
        runInfo = []
        for row in infoTab:
            if submission == str(row.get('Submission')):
                runInfo.append(row)
        if 0 == len(runInfo):
            raise Exception('Found %d entries in SRA for "%s" when expecting at least 1' % (len(runInfo), submission))
        else:
            return runInfo    
        
    @staticmethod
    def getInfoFromSampleAcc(sample):
        if not sample.startswith("ERS") and not sample.startswith("SRS"):
            raise Exception("sample should start with ERS or SRS, not: " + sample)
        infoTab = SRAUtils.getInfoTableFromSearchTerm(sample)
        runInfo = []
        for row in infoTab:
            if sample == str(row.get('Sample')):
                runInfo.append(row)
        if 0 == len(runInfo):
            raise Exception('Found %d entries in SRA for "%s" when expecting at least 1' % (len(runInfo), sample))
        else:
            return runInfo
        
    @staticmethod
    def getInfoFromProjectAcc(project):
        if not project.startswith("ERP") and not project.startswith("SRP"):
            raise Exception("project should start with ERP or SRP, not: " + project)
        infoTab = SRAUtils.getInfoTableFromSearchTerm(project)
        runInfo = []
        for row in infoTab:
            if project == str(row.get('SRAStudy')):
                runInfo.append(row)
        if 0 == len(runInfo):
            raise Exception('Found %d entries in SRA for "%s" when expecting at least 1' % (len(runInfo), project))
        else:
            return runInfo
        
    @staticmethod
    def getInfoFromExperimentAcc(experiment):
        if not experiment.startswith("ERX") and not experiment.startswith("SRX"):
            raise Exception("experiment should start with ERX or SRX, not: " + experiment)
        infoTab = SRAUtils.getInfoTableFromSearchTerm(experiment)
        runInfo = []
        for row in infoTab:
            if experiment == str(row.get('Experiment')):
                runInfo.append(row)
        if 0 == len(runInfo):
            raise Exception('Found %d entries in SRA for "%s" when expecting at least 1' % (len(runInfo), experiment))
        else:        
            return runInfo
    
    @staticmethod
    def getInfoFromBioProjectAcc(bioProject):
        # or 
        if not bioProject.startswith("PRJNA") and not bioProject.startswith("PRJEA") and not bioProject.startswith("PRJEB") and not bioProject.startswith("PRJDA"):
            raise Exception("bioProject should start with PRJNA, PRJEA, PRJEB, and PRJDA, not: " + bioProject)
        infoTab = SRAUtils.getInfoTableFromSearchTerm(bioProject)
        runInfo = []
        for row in infoTab:
            if bioProject == str(row.get('BioProject')):
                runInfo.append(row)
        if 0 == len(runInfo):
            raise Exception('Found %d entries in SRA for "%s" when expecting at least 1' % (len(runInfo), bioProject))
        else:        
            return runInfo   
    
    
        
    @staticmethod
    def getInfoFromSRAIdentifier(identifier):
        if identifier.startswith("ERX") or identifier.startswith("SRX"):
            return SRAUtils.getInfoFromExperimentAcc(identifier)
        elif identifier.startswith("ERP") or identifier.startswith("SRP"):
            return SRAUtils.getInfoFromProjectAcc(identifier)
        elif identifier.startswith("ERS") or identifier.startswith("SRS"):
            return SRAUtils.getInfoFromSampleAcc(identifier)
        elif identifier.startswith("ERR") or identifier.startswith("SRR"):
            return SRAUtils.getInfoFromRunAcc(identifier)
        elif identifier.startswith("ERA") or identifier.startswith("SRA"):
            return SRAUtils.getInfoFromSubmissionAcc(identifier)
        elif identifier.startswith("PRJDA"):
            #DDBJ archvie bioproject prefix PRJNA 
            return SRAUtils.getInfoFromBioProjectAcc(identifier)
        elif identifier.startswith("PRJNA"):
            #short read archvie bioproject prefix PRJNA
            return SRAUtils.getInfoFromBioProjectAcc(identifier)
        elif identifier.startswith("PRJEA") or identifier.startswith("PRJEB"):
            #european archive bioproject prefixes PRJEA or PRJEB
            return SRAUtils.getInfoFromBioProjectAcc(identifier)
        else:
            raise Exception("Error, unrecognized prefix for sra Identifier " + str(identifier))


def parse_args_sraIdentifier():
    parser = argparse.ArgumentParser()
    parser.add_argument('--identifiers', type=str, help = "A list of comma separated SRA identifiers e.g. SRP046206,SRX188939,SRS807544,SRR1759594,PRJNA63661",  required = True)
    parser.add_argument('--outStub', type=str, help = "An output stub for info and sra urls output files",  required = True)
    parser.add_argument('--overWrite', action = "store_true", help = "Overwrite files if they already exist")
    return parser.parse_args()

def runGetRunsFromSampleAcc():
    args = parse_args_sraIdentifier()
    identifiers = args.identifiers.split(",")
    #sys.stdout.write(str("identifier") + "\t" + str("run") + "\t" + "url" + "\n")
    outUrlsFnp = args.outStub + "_urls.tab.txt"
    outInfoFnp = args.outStub + "_info.tab.txt"
    if os.path.exists(outUrlsFnp) and not args.overWrite:
        raise Exception("File " + outUrlsFnp + " already exists, use --overWrite to over write it")
    if os.path.exists(outInfoFnp) and not args.overWrite:
        raise Exception("File " + outInfoFnp + " already exists, use --overWrite to over write it")
    with open(outUrlsFnp, "w") as outUrlsFile:
        with open(outInfoFnp, "w") as outInfoFile:
            outUrlsFile.write(str("identifier") + "\t" + str("run") + "\t" + "url" + "\n")
            identifierCount = 0
            for identifier in identifiers:
                try:
                    tab = SRAUtils.getInfoFromSRAIdentifier(identifier)
                    if 0 == identifierCount:
                        outInfoFile.write("\t".join(tab[0].keys()) + "\n")
                    for row in tab:
                        outUrlsFile.write(str(identifier) + "\t" + row.get("Run") + "\t" + SRAUtils.getSraUrlFromRunAccession(str(row.get("Run"))) + "\n")
                        outInfoFile.write("\t".join(row.values()) + "\n")
                    identifierCount = identifierCount + 1
                except Exception, err:
                    print ("Failed  to get info for " + str(identifier) + ", mess: " + str(err))
                    traceback.print_exc()
                    

if __name__ == "__main__":
    runGetRunsFromSampleAcc()



```

# Runing the script
```{r , engine='bash', eval = F}
./getSraRunsFromAccIds.py --identifiers SRP046206,SRX188939,SRS807544,SRR1759594 --outStub out
```

This will download two files, one will end with _info.tab.txt which has the whole table that querying the SRA website gets and the other one will be end with _urls.tab.txt which has the urls for all run files for the identifiers given

```{r test_download, engine='bash', eval = T, echo = FALSE}
scripts/python/getSraRunsFromAccIds.py --identifiers SRP046206,SRX188939,SRS807544,SRR1759594 --outStub out --overWrite
```


```{r test_download_report1, engine='bash', eval = T, echo = T}
cat out_info.tab.txt
```

```{r test_download_report2, engine='bash', eval = T, echo = T}
cat out_urls.tab.txt
```


#Downloading threaded 

Also below is a script that can be used on the url table created by getSraRunsFromAccIds.py above.   
This depends on the URLGrabber python library

[downloadSraFromTable.py](scripts/python/downloadSraFromTable.py)

```{r, engine='bash'}
pip install URLGrabber
```

***downloadSraFromTable.py***
```{r , engine='python', eval = F}
#!/usr/bin/env python
import shutil, os, argparse, sys, stat, urllib

from urlgrabber.grabber import URLGrabber
import requests
import csv
from joblib import Parallel, delayed
import traceback

def get_file_if_size_diff(url, d):
    fn = url.split('/')[-1]
    out_fnp = os.path.join(d, fn)
    g = URLGrabber(reget = "simple")
    locFnp = g.urlgrab(url, out_fnp)
    return locFnp
    
def downloadFileAttempt(url, directoryName):
    try:
        return get_file_if_size_diff(url, directoryName)
    except Exception, err:
        print ("Failed  to download " + str(url) + " to " + str(directoryName) + ", mess: " + str(err))
        traceback.print_exc()
        
def parse_args_downloadFiles():
    parser = argparse.ArgumentParser()
    parser.add_argument('--sraUrlFnp', type=str, required = True, help = "SRA url file created by getSraRunsFromAccIds.py")
    parser.add_argument('--outDirectory', type=str, default = "./", help = "The directory in which to download the files")
    parser.add_argument('--ncpus', type=int, default = 1, help = "Number of cpus to use")
    return parser.parse_args()   


if __name__ == "__main__":
    args = parse_args_downloadFiles()
    with open(args.sraUrlFnp, 'rb') as urlFile:
        reader =  csv.DictReader(urlFile, delimiter='\t')
        urls = []
        for row in reader:
            urls.append(row.get("url"))
        downloadedFiles = Parallel(n_jobs = args.ncpus)(delayed(downloadFileAttempt)(url, "./") for url in urls )
        print(downloadedFiles)

```

## Running the script
```{r , engine='bash', eval = F}
./downloadSraFromTable.py --sraUrlFnp out_urls.tab.txt --ncpus 6 --outDirectory ./
```

```{r test_download_of_sra, engine='bash', eval = T, echo = FALSE}
scripts/python/downloadSraFromTable.py --sraUrlFnp out_urls.tab.txt --ncpus 6 --outDirectory ./
```

The script will often time out on requests especially if you do a lot of downloads at once but simply re-run script will allow you to download all the files since it will check to see if a file is already downloaded and is the right file, so the script can be re-ran until all files have finished downloading. 

```{r , engine='bash', eval = F}
./downloadSraFromTable.py --sraUrlFnp out_urls.tab.txt --ncpus 6 --outDirectory ./
```

```{r test_download_of_sra, engine='bash', eval = T, echo = FALSE}
scripts/python/downloadSraFromTable.py --sraUrlFnp out_urls.tab.txt --ncpus 6 --outDirectory ./
```



